Assessing Ranking Fidelity of Competition Structures via Weighted Mutual Information

### 2025-10-17 CMSAC_presentation/CMSAC_reveal_js.qmd
Updated dataframe format

---
title: "Assessing Ranking Fidelity of Competition Structures via Weighted Mutual Information"
author: "Zach Culp, Josie Peterburs, Dr. Gregory J. Matthews, Dr. Ryan P.A. McShane"
format: revealjs
editor: visual
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tinytex)
library(tidyverse)
library(combinat)
load("curves_data.RData")
#load("tournament_comparison_plot.png")
#load("tournament_comparison_plot_close.png")
```

@@

```{r}
rhat_df %>% 
  select(-squared_error) %>%   
  filter(permutation %in% c("1234", "4321", "1324", "2143")) %>%   
  arrange(permutation,k) %>%   
  rename("Weighted Kendall" = weighted_kendall)
```

@@

```{r}
rhat_df %>%    
  filter (k==4) %>%   
  select(-k) %>%   
  arrange(desc(weighted_kendall)) %>%    
  rename("Weighted Kendall"= weighted_kendall, "Squared Error" = squared_error)
```

@@

-   Given the assigned "known strength", $\theta$, we simulate a game using the Bradley-Terry Model: $\Pr(i > j) = \frac{e^{\theta_i}}{e^{\theta_i} + e^{\theta_j}} = \frac{1}{1+e^{\theta_i - \theta_j}}$
-   Below is output from one simulation of a single elimination tournament structure

```{r}
renamed_df <- all_results_norm %>%    
  select(-c(simulation, distribution, seed, teams)) %>%   
  rename(theta = true_strength, r = true_rank, r_hat = rank_hat) %>%   
  mutate(theta = round(theta,3))  
head(renamed_df,8)
```

@@

```{r out.width="49%", fig.show='hold', fig.align='center'}
knitr::include_graphics("tournament_comparison_plot.png")
```

## Close-Up Results

```{r out.width="49%", fig.show='hold', fig.align='center'}
knitr::include_graphics("tournament_comparison_plot_close.png")
```

## Limitations

-   The "known" strengths are assumptions that can be estimated but never truly known
-   Ties are not addressed in any of the games
-   Ties in the $\hat{r}$ are randomly assigned

## Future Work

-   Evaluate structures using various distributions and "known" weights
-   Estimate the $\theta$s in a real-world example and compare recent changes to the old format

>>>>

---
title: "Assessing Ranking Fidelity of Competition Structures via Weighted Mutual Information"
author: "Zach Culp, Josie Peterburs, Dr. Gregory J. Matthews, Dr. Ryan P.A. McShane"
format: 
  revealjs: 
    df-print: paged
editor: visual

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tinytex)
library(tidyverse)
library(combinat)
library(knitr)
library(kableExtra)
load("curves_data.RData")
#load("tournament_comparison_plot.png")
#load("tournament_comparison_plot_close.png")
```

@@

```{r}
# Get the filtered data
table_data <- rhat_df %>% 
  select(-squared_error) %>%   
  filter(permutation %in% c("1234", "4321", "1324", "2143")) %>%   
  arrange(permutation, k) %>%   
  rename("Weighted Kendall" = weighted_kendall)

# Split into two halves
n <- nrow(table_data)
left_table <- table_data %>% head(ceiling(n/2))
right_table <- table_data %>% tail(floor(n/2))

# Create side-by-side tables
left_html <- left_table %>% 
  kable(format = "html", row.names = FALSE) %>% 
  kableExtra::kable_styling(font_size = 24)

right_html <- right_table %>% 
  kable(format = "html", row.names = FALSE) %>% 
  kableExtra::kable_styling(font_size = 24)
```
::: {.columns}
::: {.column width="48%"}
```{r}
#| echo: false
left_html
```
:::
::: {.column width="%48%"}
```{r}
#| echo: false
right_html
```
:::
:::

@@

```{r}
# Get the filtered data
table_data <- rhat_df %>%    
  filter(k == 4) %>%   
  select(-k) %>%   
  arrange(desc(weighted_kendall)) %>%    
  rename("Weighted Kendall" = weighted_kendall, "Squared Error" = squared_error)

# Split into two halves
n <- nrow(table_data)
left_table <- table_data %>% head(ceiling(n/2))
right_table <- table_data %>% tail(floor(n/2))

# Create side-by-side tables
left_html <- left_table %>% 
  kable(format = "html", row.names = FALSE) %>% 
  kableExtra::kable_styling(font_size = 20)
right_html <- right_table %>% 
  kable(format = "html", row.names = FALSE) %>% 
  kableExtra::kable_styling(font_size = 20)

```

::: {.columns}
::: {.column width="48%"}
```{r}
#| echo: false
left_html
```
:::
::: {.column width="%48%"}
```{r}
#| echo: false
right_html
```
:::
:::

@@

-   Given the assigned "known strength", $\theta$, we simulate a game using the Bradley-Terry Model: $\Pr(i > j) = \frac{e^{\theta_i}}{e^{\theta_i} + e^{\theta_j}} = \frac{1}{1+e^{\theta_i - \theta_j}}$
- Below is an example of the output of a single tournament simulation:

```{r}
renamed_df <- all_results_norm %>%    
  select(-c(simulation, distribution, seed, teams)) %>%   
  rename(theta = true_strength, r = true_rank, r_hat = rank_hat, "Wins" = game_wins, "Losses" = game_losses) %>%   
  mutate(theta = round(theta,3)) %>%
  head(8)

# Split into two halves
n <- nrow(renamed_df)
left_table <- renamed_df %>% head(ceiling(n/2))
right_table <- renamed_df %>% tail(floor(n/2))

# Create side-by-side tables
left_html <- left_table %>% 
  kable(format = "html", row.names = FALSE) %>% 
  kableExtra::kable_styling(font_size = 24)
right_html <- right_table %>% 
  kable(format = "html", row.names = FALSE) %>% 
  kableExtra::kable_styling(font_size = 24)

```

::: {.columns}
::: {.column width="48%"}
```{r}
#| echo: false
left_html
```
:::
::: {.column width="%48%"}
```{r}
#| echo: false
right_html
```
:::
:::

@@

::: {style="position: relative;"}
::: {.fade-out fragment-index="1"}
```{r out.width="60%", fig.show='hold', fig.align='center'}
knitr::include_graphics("boxed_tournament_comparison_plot.png")
```
:::

::: {.fragment .fade-in fragment-index="1" .fade-out fragment-index="2" style="position: absolute; top: 50%; left: 0; right: 0; transform: translateY(-50%);"}
```{r out.width="60%", fig.show='hold', fig.align='center'}
knitr::include_graphics("tournament_comparison_plot.png")
```
:::

::: {.fragment .fade-in fragment-index="2" style="position: absolute; top: 50%; left: 0; right: 0; transform: translateY(-50%);"}
```{r out.width="60%", fig.show='hold', fig.align='center'}
knitr::include_graphics("tournament_comparison_plot_close.png")
```
:::
:::

## Conclusion

-   These results underscore the trade-off between tournament accuracy and practical constraints such as time, cost, and entertainment value.
-   Limitations: The "known" strengths are assumptions that can be estimated but never truly known
-   Future Work: Estimate the $\theta$s in a real-world example and compare recent changes to the old format
